An Experimental Evaluation of Hybrid Vector Search (EA\&B)

### 2025-04-28 VLDB2026/main.tex
modify

	\setlength{\textfloatsep}{0cm}
	\setlength{\floatsep}{0cm}
	\renewcommand{\arraystretch}{0.9}
	\begin{table*}[t]
		\centering
		%\setlength{\belowcaptionskip}{-0.3cm}
		\setlength{\abovecaptionskip}{0cm}
		\setlength{\belowcaptionskip}{-0.3cm}
		\caption{Comparison of AF-ANN search algorithms}
		\small
		\label{tab:compair_1}
		\begin{tabular}{|l|l|c|c|c|c|c|c|c|}
			\hline
			\textbf{Algorithm} & \textbf{Base} & \textbf{Filter Type} & \textbf{AND} & \textbf{OR} & \textbf{Flexible Attributes} & \textbf{Complex Boolean} & \textbf{Dynamic Insert} & \textbf{Multi Thread} \\
			\hline
			NHQ & Graph & C & Y & N& N& N & N & N \\
			FilteredVamana & Graph & C & N & Y & Y & N & Y & Y \\
			StitchedVamana & Graph & C & N & Y & Y & N & N & Y  \\
			CAPS & IVF & C & Y & N & Y & N & Y & Y \\
			ACORN & Graph & B & Y & Y & Y & Y & Y & Y \\
			UNG & Graph & B & Y & Y & Y & N & Y & Y \\ 
			Puck & IVF & B & Y & Y & Y & N & Y & Y \\
			% ParlayANN & Graph+IVF & B & N & N & Y & N & N & Y \\

@@

		\centering
		%	 \setlength{\abovecaptionskip}{0cm}
		%	\setlength{\belowcaptionskip}{0.1cm}
		\footnotesize{
			\begin{minipage}{\linewidth}
				\vspace{0.1cm}
				\textsuperscript{A} Post-filtering, 
				\textsuperscript{B} Pre-filtering, 
				\textsuperscript{C} Simultaneous filtering, 
				\textsuperscript{Y} Support, 
				\textsuperscript{N} Unsupport. 
				\textit{Base} indicates that the hybrid query algorithm is an improvement based on a specific ANN search algorithm. 
				\textit{AND} indicates whether the algorithm supports attribute filtering with AND operations. 
				\textit{OR} indicates whether the algorithm supports attribute filtering with OR operations. 
				\textit{Flexible Attributes} indicates whether the number of query attributes can differ from the number of attributes used during index construction. 
				\textit{Complex Boolean} refers to whether the algorithm supports advanced Boolean logic beyond just AND and OR, such as NOT or combinations like (A AND B) OR (C AND NOT D). 
				\textit{Dynamic} indicates whether the algorithm supports the operation of dynamically inserting data points.
		\end{minipage}}
		\vspace{-0.5cm}
	\end{table*}

@@

	\noindent\textbf{\underline{Filtered-DiskANN \cite{Filtered-diskann}.}}  
	\sout{The effectiveness of NHQ may degrade when the number of attributes changes, as it affects the fusion distance computation. Filtered-DiskANN addresses this limitation.}
	Filtered-DiskANN also supports simultaneous filtering. Built upon the Vamana \cite{diskann} graph-based ANN index, it incorporates attribute information directly into the graph during index construction. This integration ensures that the index reflects both vector similarity and attribute constraints. Moreover, Filtered-DiskANN supports SSD-based storage, improving scalability.
	Filtered-DiskANN proposes 2 indexes:  
	1) \textbf{FilteredVamana}, which incrementally builds the graph index by inserting data points and dynamically adding edges, allowing adaptive expansion.  
	2) \textbf{StitchedVamana}, which adopts a batch construction strategy—constructing separate Vamana subgraphs for each attribute, followed by merging and edge pruning.
	A key limitation of Filtered-DiskANN is that while it supports both single-attribute and multi-attribute filtering, it only supports the Boolean OR logic for multi-attribute filtering, lacking support for Boolean AND logic.

	\noindent\textbf{\underline{CAPS \cite{CAPS}.}} 
	\sout{Different from the above graph-based simultaneous filtering algorithms,}CAPS is the first simultaneous filtering algorithm based on spatial partitioning.
	%As graph-based methods, NHQ and Filtered-DiskANN often result in large index sizes. CAPS addresses this issue by being the first simultaneous filtering algorithm based on spatial partitioning, significantly reducing index size.
	CAPS introduces a hierarchical sub-partitioning algorithm inspired by Huffman trees, termed the Attribute Frequency Tree (AFT), to overcome the coarse granularity of traditional IVF-based methods. CAPS adopts a two-level partitioning strategy:  
	1) The first level clusters vectors based on similarity using K-Means or learning-based methods such as BLISS. 2) Within each cluster, AFT partitions data further based on attribute frequencies, enabling finer-grained indexing and improving query efficiency.

@@

		\footnotesize{
			\begin{minipage}{\linewidth}
				\textsuperscript{Y} Support, 
				\textsuperscript{N} Unsupport.
				\textit{Base} indicates that the hybrid query algorithm is an improvement based on a specific ANN search algorithm. 
				\textit{Dynamic} indicates whether the algorithm supports the operation of dynamically inserting data points.
			\end{minipage} 
			\vspace{-0.1cm}
		}

@@

	\begin{figure*}
		\centering
		\setlength{\abovecaptionskip}{0cm}
		\setlength{\belowcaptionskip}{-0.5cm}
		\includegraphics[width=0.96\textwidth]{figures/exp/exp_1_1_SingleLabel_1thread.pdf}
		\caption{Single-Attribute Building and Query (single thread) }
		\label{fig:exp_1_1_SingleLabel_1thread}
	\end{figure*}

@@

	\textit{\textbf{Index size.}}
	Traditional graph-based methods often only evaluate the size of the generated graph index and do not include the original data, while the index generated by the IVF-based method includes the original data. In order to unify the comparison standard, 
	we include both the index structure generated by the algorithm and the original data when evaluating the index size. 
	% Traditional graph-based methods often evaluate only the size of the generated index, excluding the original data, whereas IVF-based methods include it. To ensure a consistent comparison, we consider both the index structure and the original data when measuring index size.
	As shown in the Figure~\ref{fig:index_size_mb_comparison_query1}, PASE exhibits significantly larger index sizes across all supported datasets. This overhead stems primarily from its extensive metadata requirements. The index sizes of other methods show moderate variation. Among these, Faiss generates slightly smaller indexes due to its efficient index design.

@@

	\begin{figure*}
		\centering
		\setlength{\abovecaptionskip}{0.0cm}
		\setlength{\belowcaptionskip}{-0.4cm}
		\includegraphics[width=0.95\textwidth]{figures/exp/exp_4_1_MultiLabel_1thread.pdf}
		\caption{Multi-attribute build and query (single thread)}
		\label{fig:exp_4_1_MultiLabel_1thread}
	\end{figure*}

	\begin{figure*}
		\centering
		\setlength{\abovecaptionskip}{0.0cm}
		\setlength{\belowcaptionskip}{-0.4cm}
		\includegraphics[width=0.95\textwidth]{figures/exp/exp_3_1.pdf}
		\caption{Effect of attribute distribution on query performance (single thread)}
		\label{fig:exp_3_1}
	\end{figure*}

@@

	\begin{figure}
		\centering
		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.1cm}
		\includegraphics[width=0.95\columnwidth]{figures/exp/exp_5_1_1_SingleLabel_1thread.pdf}
		\caption{Effect of Single-Attribute Selectivity on Query}
		\label{fig:exp_5_1_1_SingleLabel_1thread}
	\end{figure}

@@

	\begin{figure*}
		\centering
		    \setlength{\abovecaptionskip}{-0.1cm}
		    \setlength{\belowcaptionskip}{-0.4cm}
		\includegraphics[width=0.97\textwidth]{figures/exp/exp_6_1.pdf}
		\caption{Effect of dataset on Query}
		\label{fig:exp_6_1}
	\end{figure*}

@@

	\begin{figure*}[t]

		\centering
		\setlength{\abovecaptionskip}{0cm}
		\setlength{\belowcaptionskip}{-0.5cm}
		\includegraphics[width=0.92\textwidth]{figures/exp/exp_8_2.pdf}
		\caption{RF-ANN search }
		\label{fig:exp_8_2}
	\end{figure*}

@@

				\hline
				S5: General-purpose Use & Milvus, Faiss &DSG, UNIFY\\
				\hline
				S6: Resource constraint& CAPS, Faiss&Faiss, SeRF \\
				\hline
				S7: High-Recall Querying & UNG & WinFilter \\
				\hline
				S8: OOD scenarios & UNG &WinFilter  \\
				\hline
			\end{tabular}
		}
		\label{tab:algo_scenarios}
	\vspace{-0.0cm}
	\end{table}

>>>>

	\setlength{\textfloatsep}{0cm}
	\setlength{\floatsep}{0cm}
	\renewcommand{\arraystretch}{0.9}
	\begin{table*}[t]
		\centering
		%\setlength{\belowcaptionskip}{-0.3cm}
		\setlength{\abovecaptionskip}{0cm}
		\setlength{\belowcaptionskip}{-0.3cm}
		\caption{Comparison of AF-ANN search algorithms}
		\small
		\label{tab:compair_1}
		\begin{tabular}{|l|l|c|c|c|c|c|c|c|}
			\hline
			\textbf{Algorithm} & \textbf{Base} & \textbf{Filter Type} & \textbf{AND} & \textbf{OR} & \textbf{Flexible Attributes} & \textbf{Complex Boolean} & \textbf{Dynamic Insert} & \textbf{Multi Thread} \\
			\hline
			NHQ & Graph & Simultaneous filtering & Y & N& N& N & N & N \\
			FilteredVamana & Graph & Simultaneous filtering & N & Y & Y & N & Y & Y \\
			StitchedVamana & Graph & Simultaneous filtering & N & Y & Y & N & N & Y  \\
			CAPS & IVF &Simultaneous filtering & Y & N & Y & N & Y & Y \\
			ACORN & Graph & Pre-filtering & Y & Y & Y & Y & Y & Y \\
			UNG & Graph & Pre-filtering & Y & Y & Y & N & Y & Y \\ 
			Puck & IVF & Pre-filtering & Y & Y & Y & N & Y & Y \\
			% ParlayANN & Graph+IVF & B & N & N & Y & N & N & Y \\

@@

		\centering
		%	 \setlength{\abovecaptionskip}{0cm}
		%	\setlength{\belowcaptionskip}{0.1cm}
		\footnotesize{
			\begin{minipage}{\linewidth}
				\vspace{0.1cm}
%				\textsuperscript{A} Post-filtering, 
%				\textsuperscript{B} Pre-filtering, 
%				\textsuperscript{C} Simultaneous filtering, 
				\textsuperscript{Y} Support, 
				\textsuperscript{N} Unsupport. 
				\textit{Base} indicates that the hybrid query algorithm is an improvement based on a specific ANN search algorithm. 
				\textit{AND} indicates whether the algorithm supports attribute filtering with AND operations. 
				\textit{OR} indicates whether the algorithm supports attribute filtering with OR operations. 
				\textit{Flexible Attributes} indicates whether the number of query attributes can differ from the number of attributes used during index construction. 
				\textit{Complex Boolean} refers to whether the algorithm supports advanced Boolean logic beyond just AND and OR, such as NOT or combinations like (A AND B) OR (C AND NOT D). 
				\textit{Dynamic} indicates whether the algorithm supports the operation of dynamically inserting data points.
				\textit{Multi Thread} indicates whether the algorithm supports multi-threaded search.
		\end{minipage}}
		\vspace{-0.5cm}
	\end{table*}

@@

	\noindent\textbf{\underline{Filtered-DiskANN \cite{Filtered-diskann}.}}  
%	\sout{The effectiveness of NHQ may degrade when the number of attributes changes, as it affects the fusion distance computation. Filtered-DiskANN addresses this limitation.}
	Filtered-DiskANN also supports simultaneous filtering. Built upon the Vamana \cite{diskann} graph-based ANN index, it incorporates attribute information directly into the graph during index construction. This integration ensures that the index reflects both vector similarity and attribute constraints. Moreover, Filtered-DiskANN supports SSD-based storage, improving scalability.
	Filtered-DiskANN proposes 2 indexes:  
	1) \textbf{FilteredVamana}, which incrementally builds the graph index by inserting data points and dynamically adding edges, allowing adaptive expansion.  
	2) \textbf{StitchedVamana}, which adopts a batch construction strategy—constructing separate Vamana subgraphs for each attribute, followed by merging and edge pruning.
	A key limitation of Filtered-DiskANN is that while it supports both single-attribute and multi-attribute filtering, it only supports the Boolean OR logic for multi-attribute filtering, lacking support for Boolean AND logic.

	\noindent\textbf{\underline{CAPS \cite{CAPS}.}} 
	%\sout{Different from the above graph-based simultaneous filtering algorithms,}
	CAPS is the first simultaneous filtering algorithm based on spatial partitioning.
	%As graph-based methods, NHQ and Filtered-DiskANN often result in large index sizes. CAPS addresses this issue by being the first simultaneous filtering algorithm based on spatial partitioning, significantly reducing index size.
	CAPS introduces a hierarchical sub-partitioning algorithm inspired by Huffman trees, termed the Attribute Frequency Tree (AFT), to overcome the coarse granularity of traditional IVF-based methods. CAPS adopts a two-level partitioning strategy:  
	1) The first level clusters vectors based on similarity using K-Means or learning-based methods such as BLISS. 2) Within each cluster, AFT partitions data further based on attribute frequencies, enabling finer-grained indexing and improving query efficiency.

@@

		\footnotesize{
			\begin{minipage}{\linewidth}
				\centering
				\textcolor{red}{The evaluation metrics are consistent with those in Table ~\ref{tab:compair_1}.}
%				\textsuperscript{Y} Support, 
%				\textsuperscript{N} Unsupport.
%				\textit{Base} indicates that the hybrid query algorithm is an improvement based on a specific ANN search algorithm. 
%				\textit{Dynamic} indicates whether the algorithm supports the operation of dynamically inserting data points.
			\end{minipage} 
			\vspace{-0.1cm}
		}

@@

	\begin{figure*}
		\centering
		\setlength{\abovecaptionskip}{0cm}
		\setlength{\belowcaptionskip}{-0.5cm}
		\includegraphics[width=0.98\textwidth]{figures/exp/exp_1_1_SingleLabel_1thread.pdf}
		\caption{Single-Attribute Building and Query (single thread) }
		\label{fig:exp_1_1_SingleLabel_1thread}
	\end{figure*}

@@

	\textit{\textbf{Index size.}}
	\sout{Traditional graph-based methods often only evaluate the size of the generated graph index and do not include the original data, while the index generated by the IVF-based method includes the original data. }
	In order to unify the comparison standard, we include both the index structure generated by the algorithm and the original data when evaluating the index size. 
	% Traditional graph-based methods often evaluate only the size of the generated index, excluding the original data, whereas IVF-based methods include it. To ensure a consistent comparison, we consider both the index structure and the original data when measuring index size.
	As shown in the Figure~\ref{fig:index_size_mb_comparison_query1}, PASE exhibits significantly larger index sizes across all supported datasets. This overhead stems primarily from its extensive metadata requirements. The index sizes of other methods show moderate variation. Among these, Faiss generates slightly smaller indexes due to its efficient index design.

@@

	\begin{figure*}
		\centering
		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.4cm}
		\includegraphics[width=\textwidth]{figures/exp/exp_4_1_MultiLabel_1thread.pdf}
		\caption{Multi-attribute build and query (single thread)}
		\label{fig:exp_4_1_MultiLabel_1thread}
	\end{figure*}

	\begin{figure*}
		\centering
		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.4cm}
		\includegraphics[width=\textwidth]{figures/exp/exp_3_1.pdf}
		\caption{Effect of attribute distribution on query performance (single thread)}
		\label{fig:exp_3_1}
	\end{figure*}

@@

	\begin{figure}
		\centering
		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.1cm}
		\includegraphics[width=\columnwidth]{figures/exp/exp_5_1_1_SingleLabel_1thread.pdf}
		\caption{Effect of Single-Attribute Selectivity on Query}
		\label{fig:exp_5_1_1_SingleLabel_1thread}
	\end{figure}

@@

	\begin{figure*}
		\centering
		    \setlength{\abovecaptionskip}{0cm}
		    \setlength{\belowcaptionskip}{-0.4cm}
		\includegraphics[width=0.97\textwidth]{figures/exp/exp_6_1.pdf}
		\caption{Effect of dataset on Query}
		\label{fig:exp_6_1}
	\end{figure*}

@@

	\begin{figure*}[h]
		\centering
		\setlength{\abovecaptionskip}{0cm}
		\setlength{\belowcaptionskip}{-0.5cm}
		\includegraphics[width=0.95\textwidth]{figures/exp/exp_8_2.pdf}
		\caption{RF-ANN search }
		\label{fig:exp_8_2}
	\end{figure*}

@@

				\hline
				S5: General-purpose Use & Milvus, Faiss &DSG, UNIFY\\
				\hline
				S6: Resource constraint& CAPS, Faiss&Faiss, SeRF \\
				\hline
				S7: High-Recall Querying & UNG & WinFilter \\
				\hline
				S8: OOD scenarios & UNG &WinFilter  \\
				\hline
			\end{tabular}
		}
		\label{tab:algo_scenarios}
	\vspace{0cm}
	\end{table}
